Moment Distances from Robust Subspace for Botnet Detection

### 2019-05-14 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
mdrs

The general formula of the $r$-th moment can be expressed as:
\begin{equation}\label{eq:eq02}
	m_r = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^r. 
\end{equation}

Therefore, the Skewness is calculated as
\begin{equation}\label{eq:eq03}
	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
\end{equation}

and Kurtosis as
\begin{equation}\label{eq:eq04}
	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2} .
\end{equation}

We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq01} for $\textbf{X}^r$. We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies betweem new observations and moments computed from a robust subspace computed by RPCA. The classical equation of we modify

The classifical MD relies on robust mean and robust covariace matrix, here our contribution is compute the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA, instead of classifical approach, that relies on MCD or Fast-MCD.

We propose to modify the Equation \ref{eq:eq01} to implement a Skewness-based Mahalanobis Distance, as follows:

\begin{equation}\label{eq:eq08}
	\boldsymbol{d}(\boldsymbol{s}_{\boldsymbol{x}}, \bar{\boldsymbol{s}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{s}_{\boldsymbol{x}} - \bar{\boldsymbol{s}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{s}_{\boldsymbol{x}} - \bar{\boldsymbol{s}})^\prime}, 
\end{equation}

where $\boldsymbol{s}_{\boldsymbol{x}} = \boldsymbol{x} - \boldsymbol{\bar{s}}(\boldsymbol{x})$ and $\boldsymbol{\bar{s}}(\boldsymbol{x})$ is the Skewness of $\boldsymbol{x}$.

We also propose to modify the Equation \ref{eq:eq04} to implement a Kurtosis-based Mahalanobis Distance, as follows:

\begin{equation}\label{eq:eq09}
	\boldsymbol{d}(\boldsymbol{k}_{\boldsymbol{x}}, \bar{\boldsymbol{k}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{k}_{\boldsymbol{x}} - \bar{\boldsymbol{k}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{k}_{\boldsymbol{x}} - \bar{\boldsymbol{k}})^\prime}, 
\end{equation}

where $\boldsymbol{k}_{\boldsymbol{x}} = \boldsymbol{x} - \boldsymbol{\bar{k}}(\boldsymbol{x})$ and $\boldsymbol{\bar{k}}(\boldsymbol{x})$ is the Kurtosis of $\boldsymbol{x}$.

% \begin{algorithm}
% 	\label{alg:alg01}
% 	\scriptsize
% 	\SetAlgoLined
% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$}
% 	Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
% 	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
% 		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
% 			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
% 		}
% 		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
% 		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
% 		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
% 		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
% 		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
% 	}
% 	Compute $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$ for last $\boldsymbol{h}_2$\;
% 	\caption{C-Step with higher moments}
% \end{algorithm}

@@

\subsection{MOS Schemes}
\label{sec:prop_MOSSchemes}

\subsection{Eigen Similarity Analysis}
\label{sec:prop_EigenSimilarityAnalysis}

\subsubsection{Time Similarity Analysis}
\label{sec:prop_TimeSimilarityAnalysis}

\subsubsection{Port Similarity Analysis}
\label{sec:prop_PortSimilarityAnalysis}

@@

% Mahalanobis Distance is defined as		
% 	\begin{equation}\label{eq:eq01}
% 		d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}.
% 	\end{equation}
% $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector, also referred as location, of known observations and $\boldsymbol{\hat{S}}$ is the covariance matrix, also referred as scatter, of known observations.

@@

% Let the mean $\bar{\boldsymbol{x}} \in \mathbb{R}^{1 \times p}$ be

% \begin{equation}\label{eq:eq02}
% 	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} \boldsymbol{x}_j, 
% \end{equation}

% the covariance matrix $\boldsymbol{\hat{S}} \in \mathbb{R}^{p \times p}$ be

% \begin{equation}\label{eq:eq03}
% 	\boldsymbol{\hat{S}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} (\boldsymbol{x}_j - \bar{\boldsymbol{x}})(\boldsymbol{x}_j - \bar{\boldsymbol{x}})^\prime,
% \end{equation}

>>>>

Let the mean $\bar{\boldsymbol{x}} \in \mathbb{R}^{1 \times p}$ be

\begin{equation}\label{eq:eq02}
	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} \boldsymbol{x}_i, 
\end{equation}

and the covariance matrix $\boldsymbol{\bar{\Sigma}} \in \mathbb{R}^{p \times p}$ be

\begin{equation}\label{eq:eq03}
	\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime,
\end{equation}

The general formula of the $u$-th moment can be expressed as
\begin{equation}\label{eq:eq02}
	m_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^u.
\end{equation}

Therefore, the skewness is calculated by
\begin{equation}\label{eq:eq03}
	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
\end{equation}

and the kurtosis as
\begin{equation}\label{eq:eq04}
	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2} .
\end{equation}

We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq01} for $\textbf{X}^r$. We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies betweem new observations and the moments calculated from a robust subspace computed by RPCA.

The classical Mahalanobis Distance is defined as		
	\begin{equation}\label{eq:eq05}
		\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime},
	\end{equation}
where $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector of known observations, also referred as location, and $\boldsymbol{\bar{\Sigma}}$ is the covariance matrix of known observations, also referred as scatter. The classifical MD is usually relies on robust mean and robust covariace matrix for outlier detection, which are commonly computed by Fast-MCD \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}. Here we propose to calculate the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA. We also propose to extend the Equation \ref{eq:eq05} to implement a Skewness-based Mahalanobis Distance, as follows:

\begin{equation}\label{eq:eq08}
	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}.
\end{equation}

We finally propose to extend the Equation \ref{eq:eq05} to implement a Kurtosis-based Mahalanobis Distance, as follows:

\begin{equation}\label{eq:eq09}
	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}.
\end{equation}

The distances shall be computed and evaluated separatelly for botnet detection, in order to verify if the scenarios have different patterns that can be explored differently by distances of each moment individually.

The Algorithmm \ref{alg:alg01} presents a overview the experiment
\begin{algorithm}
	\label{alg:alg01}
	\scriptsize
	\SetAlgoLined
	\KwResult{$\boldsymbol{\bar{t}}$}
	Given the data set $\boldsymbol{X}$ split into 13 scenarios $\boldsymbol{X}_i$, and susequently split into normal observations $\boldsymbol{X}_i^n$ and contaminated observations $\boldsymbol{X}_i^t$\;
	\For{$eacch scenario$}{
		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
		}
		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
	}
	Compute $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$ for last $\boldsymbol{h}_2$\;
	\caption{C-Step with higher moments}
\end{algorithm}

The expected result $\boldsymbol{\bar{t}}$ refers to the predicted classification of the new observations as normal or anomalous.

@@

@@

@@
